---
title: "Food Insecurity: Violent Development in Sub-Saharan Africa"
author: Bill Foote
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include = FALSE, eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE}

set.seed(4289) #as always, the answer is 42, or is it 89
library(lubridate)
library(RColorBrewer)
library(tidyverse)
library(kableExtra)
library(plotly)
#
pop <- read_csv( "population.csv")
pop_long <- pop |>
  gather( key=year, value=population, '1990':'2021')
# ggplot helper
# 
plot_2_grid <- function( plot1, plot2, title_1= "Default" ){ 
    # build side by side plots
    plot_row <- plot_grid(plot1, plot2)
    # now add the title
    title <- ggdraw() + 
    draw_label(
      title_1,
      fontface = 'bold',
      x = 0,
      hjust = 0
    ) +
    theme(
      # add margin on the left of the drawing canvas,
      # so title is aligned with left edge of first plot
      plot.margin = margin(0, 0, 0, 7)
    )
    plot_grid(
    title, plot_row,
    ncol = 1,
    # rel_heights values control vertical title margins
    rel_heights = c(0.1, 1)
    )
  }
plot_3_grid <- function( plot1, plot2, plot3, title_1= "Default" ){ 
    # build side by side plots
    plot_row <- plot_grid(plot1, plot2, plot3)
    # now add the title
    title <- ggdraw() + 
    draw_label(
      title_1,
      fontface = 'bold',
      x = 0,
      hjust = 0
    ) +
    theme(
      # add margin on the left of the drawing canvas,
      # so title is aligned with left edge of first plot
      plot.margin = margin(0, 0, 0, 7)
    )
    plot_grid(
    title, plot_row,
    ncol = 1,
    # rel_heights values control vertical title margins
    rel_heights = c(0.1, 1)
    )
}
#
# load data
#vio <- read_csv( "gpd-violence-sub-sahara.csv")
d_all <- read_csv( "casualties-life-gdppc-map-0.csv" )
```

Sourcing Food Insecurity: Violent Development
===========================


Column {.tabset}
---------------------------

### Tracking violence

```{r}
df <- d_all #|> 
#  filter( year==2018)
#df$q <- with(df, cut(casualties, quantile(casualties)))
#levels(df$q) <- paste(c("1st", "2nd", "3rd", "4th", "5th"), "Quantile")
#df$q <- as.ordered(df$q)

g <- list(
  scope = 'africa',
  showland = TRUE,
  #showframe = TRUE,
  landcolor = toRGB("gray85"),
  subunitcolor = toRGB("white"),
  countrycolor = toRGB("white")
)

fig <- plot_geo(df, locationmode = 'country names', sizes = c(1, 600))
fig <- fig |> add_markers(
    x = ~lng, y = ~lat, size = ~casualties, color = ~casualties, frame = ~year, hoverinfo = "text",
    text = ~paste(df$year, "<br />", df$country, "<br />", "GDP pc", df$gdppc, "<br />", "Life expected", df$life, "<br />",  "Casualties", df$casualties)
  )
fig <- fig |> layout(title = 'Development and Violence in Sub-Saharan Africa', geo = g)

fig <- fig |>
  animation_opts(
    1000, easing = "elastic", redraw = TRUE
  )
fig
```

### Tracking development

```{r}
df <- d_all |> 
  filter( country != "South Africa")
fig <- df |>
  plot_ly(
    y = ~life, 
    x = ~gdppc, 
    size = ~casualties, 
    color = ~country, 
    frame = ~year, 
    text = ~country, 
    hoverinfo = "text",
    type = 'scatter',
    mode = 'markers'
  )
# only top ten each year showing
fig <- fig |>
  animation_opts(
    1000, easing = "elastic", redraw = TRUE
  )
fig
 
```

### Notes

```{css, echo=FALSE}
p {
  font-size: 18px;
}
li {
  font-size: 18px;
}
```

Many to-do's:

1. Update gdppc data and merges to most recently available which as of today is only 2019.

2. Find replacement for gdppc or at least proxy fdor >2019. Perhaps fdi as % of GDP, life satisfaction, mortality < 5 years of age, access to food, sanitation.

3. inner_join( d_all, food_inflation, fdi )

4. Investigate changing violent incident bubble size.

5. Use xlim() to modify scatterplot. Check Equatorial Africa gdppc which looks way too big.

6. Here is a good plotly online reference: <https://plotly-r.com/index.html>

### Files

Data, Rmarkdown, and HTML files are in this Google Drive directory: <https://drive.google.com/drive/folders/1-lQ8jjAp5tmDpKmCEHappn67RDyTxuOv?usp=share_link>

- Animation file: `casualties-life-gdppc-map-0.csv`

- ACLED violence: `1997-2022-violence-middle-south-east-africa.csv`

- World inflation: `Inflation-data-world.xlsx`

- Food inflation by country and year: `food_long.csv`

- FDI as \% of GDP by country and year:
`fdi_long.csv`
